% LEON: LLM-Enabled Intent-Driven Operations for LEO Mega-Constellations
% Target: IEEE INFOCOM 2027
\documentclass[10pt, conference, letterpaper]{IEEEtran}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{cite}
\usepackage{url}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{adjustbox}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, calc}

% Semantic macro for validator decisions (avoids \textsc font issues)
\newcommand{\decision}[1]{\textsf{\MakeUppercase{#1}}}

\usepackage{amsthm}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}

\begin{document}

\title{LEON: LLM-Enabled Intent-Driven Operations\\for LEO Mega-Constellations}

\author{\IEEEauthorblockN{Yuanhang Li}}

\maketitle

% ============================================================
\begin{abstract}
Large LEO mega-constellations already carry diverse services---remote
sensing data relay, telemetry tracking and command (TT\&C), and
latency-sensitive financial traffic---yet translating operator
requirements into routing constraints remains largely manual.
Three problems grow acute as constellation scale increases.
First, operator intents that combine multiple conditions
(latency bounds, region avoidance, node maintenance) are difficult
to parse correctly with rule-based methods, which degrade sharply
on compositional requests.
Second, an incorrectly compiled constraint can silently disable
critical links or violate safety margins, risking service outages
for mission-critical traffic.
Third, when disruptions such as solar storms or equipment failures
demand rapid reconfiguration, route computation on the constrained
topology must be fast enough for real-time response.

We present \textbf{LEON} (LLM-Enabled Intent-Driven Operations for
LEO Networks), an intent-driven routing framework that addresses
these challenges with three matched components.
An LLM-based intent compiler built on Qwen3.5-9B converts
natural-language operator requests into a typed constraint
intermediate representation, achieving 86.2\% full semantic match
on a 240-intent benchmark and outperforming rule-based parsing
by 59.0 percentage points on compositional intents.
A deterministic 8-pass validator checks structural correctness,
semantic consistency, and physical feasibility before any constraint
reaches the routing layer, accepting no unsafe intents across all
evaluation scenarios.
A lightweight GNN cost-to-go router executes validated constraints
on a Walker Delta 20$\times$20 constellation, delivering 99.8\%
packet delivery ratio with 17$\times$ inference speedup over
Dijkstra.
End-to-end experiments across four constrained scenarios---node
failure, plane maintenance, polar avoidance, and compositional
constraints---confirm zero constraint violations.
\end{abstract}

\begin{IEEEkeywords}
LEO satellite networks, intent-based networking, graph neural networks,
large language models, constrained routing, formal verification
\end{IEEEkeywords}

% ============================================================
\section{Introduction}
\label{sec:intro}

Large low Earth orbit (LEO) constellations are becoming critical
infrastructure for global connectivity, remote sensing backhaul,
and resilient space information services~\cite{leo_constellation_survey}.
As these systems scale, network operation extends well beyond
computing routes over a known topology. Operators must frequently
translate mission-level requirements into concrete routing actions
under time pressure. A remote sensing task may require bulk image
data to be relayed around congested regions before a downlink window
closes; telemetry, tracking, and command (TT\&C) traffic may need
protected low-latency paths during critical maneuvers; and emergency
conditions such as solar storms or equipment failures may demand
immediate rerouting away from affected inter-satellite links (ISLs).
Conventional configuration workflows---manually written policies and
static routing rules---can handle these cases individually, but they
become increasingly difficult to adapt quickly and consistently as
constellation size, service diversity, and failure modes
grow~\cite{satnet_ops}.

This growing operational complexity motivates higher-level control
abstractions. Intent-based networking (IBN) has been widely studied
in terrestrial networks as a way to let operators specify \emph{what}
they want while the system determines \emph{how} to enforce
it~\cite{ibn_survey,ibn_framework}. In parallel, satellite networking
research has investigated software-defined control and dynamic routing
for non-terrestrial networks~\cite{satellite_sdn_survey,leo_routing_survey}.
These two lines of work are complementary but do not directly solve
the operational problem above: terrestrial IBN methods assume
comparatively stable topologies and do not model orbital
dynamics~\cite{ibn_template,ibn_ontology}, while satellite control
frameworks typically require structured policy inputs rather than
free-form requests that must be compiled, checked, and executed
safely end to end.

This paper studies that missing step: converting operator intent into
safe and efficient routing actions for large LEO constellations. We
focus on a Walker Delta $20\times20$ constellation (400~satellites,
550\,km altitude, 53\textdegree\ inclination) and identify three
coupled challenges:

\begin{enumerate}[leftmargin=*]
\item \textbf{Semantic translation.} Operational requests naturally
combine multiple constraints in one statement---traffic class,
excluded regions, latency budgets, failover conditions---especially
when an operator must simultaneously prioritize TT\&C while avoiding
storm-exposed polar links, or redirect remote sensing relay traffic
while preserving path diversity. Rule-based parsers handle simple
single-constraint intents but degrade sharply on such compositional
requests (40.0\% vs.\ 99.0\% accuracy in our evaluation).

\item \textbf{Safety and feasibility.} In satellite networks, an
incorrect or infeasible policy is not merely an optimization error.
A mistranslated intent can route TT\&C traffic onto unstable paths,
delay time-sensitive sensing data, or prohibit links that are
topologically necessary after failures. Accepting an infeasible
policy can create routing blackholes or disconnect critical traffic
classes. Safety therefore requires that invalid or operationally
dangerous intents be kept off the certified execution path---either
explicitly rejected or deferred to conservative fallback routing.

\item \textbf{Execution efficiency.} Even after compilation and
verification, the network must produce routing decisions quickly
enough for real-time response. Exact shortest-path procedures deliver
high-quality routes but become a bottleneck when operators need rapid
reconfiguration across a 400-node time-varying constellation during
solar-storm exposure or satellite failures.
\end{enumerate}

To address these challenges, we present \textbf{LEON}
(\textbf{L}LM-\textbf{E}nabled Intent-driven \textbf{O}perations
for LEO \textbf{N}etworks), an end-to-end framework for safe
intent-driven routing in large satellite constellations
(Fig.~\ref{fig:system_architecture}). LEON makes
the challenge-to-solution mapping explicit:
\begin{itemize}[leftmargin=*]
\item For \emph{semantic translation}, LEON uses an LLM-based intent
compiler (Qwen3.5-9B) that converts natural-language operator
requests into a typed constraint intermediate representation,
achieving 86.2\% full semantic match and outperforming rule-based
parsing by 59.0 percentage points on compositional intents.
\item For \emph{safety and feasibility}, LEON introduces a
deterministic 8-pass validator that checks structural correctness,
semantic consistency, and constructive feasibility before any
constraint reaches the routing layer, accepting no unsafe intents
in our evaluation.
\item For \emph{execution efficiency}, LEON employs a lightweight
GNN cost-to-go router (152K parameters) that approximates
Dijkstra-quality routing with 17$\times$ inference speedup,
delivering 99.8\% packet delivery ratio.
\end{itemize}

\noindent End-to-end evaluation across four constrained
scenarios---node failure, plane maintenance, polar avoidance, and
compositional constraints---confirms zero constraint violations with
both GNN and heuristic Dijkstra fallback routing. The LLM
compiler further retains 81.8\% accuracy on 33 scorable
out-of-distribution paraphrases (4.4pp degradation), demonstrating
robustness to phrasing variation.

Our contributions are:
\begin{itemize}[leftmargin=*]
\item We formulate safe intent-driven operation for large LEO
satellite networks and present LEON, an end-to-end architecture
connecting natural-language intent compilation, deterministic safety
validation, and fast constrained routing. Unlike terrestrial IBN
systems~\cite{ibn_survey,ibn_framework} and prior satellite control
frameworks~\cite{satellite_sdn_survey,leo_routing_survey}, LEON is
designed specifically for dynamic constellation topology and
mission-critical traffic.
\item We design an LLM-based intent compiler that captures
compositional operator requests with 86.2\% full semantic match,
59.0pp improvement over rule-based parsing on compositional cases,
and 81.8\% accuracy on out-of-distribution paraphrases.
\item We develop a deterministic 8-pass validator that blocks
structurally invalid, semantically inconsistent, or infeasible
intents from certified acceptance (via explicit rejection or
abstention to conservative fallback routing), achieving 0\% unsafe acceptance and
grounding safety in operationally meaningful satellite-network
constraints.
\item We introduce a compact GNN cost-to-go router for efficient
constrained routing, achieving 99.8\% packet delivery ratio and
17$\times$ speedup, while the full LEON pipeline maintains zero
constraint violations across four operational scenarios.
\end{itemize}

The rest of this paper is organized as follows.
Section~\ref{sec:related} reviews related work.
Section~\ref{sec:system} presents the system model and overall LEON
architecture. Section~\ref{sec:eval} reports experimental
results. Section~\ref{sec:discussion} discusses design trade-offs
and limitations, and Section~\ref{sec:conclusion} concludes the paper.

% ============================================================
\section{Related Work}
\label{sec:related}

To position LEON with respect to prior literature, we review the
lines of work most relevant to this paper. Our goal is not only to
summarize existing methods, but also to clarify what they do well,
where they fall short, and why an end-to-end framework for
intent-driven LEO operation is still needed. We organize the
discussion into five categories: LEO constellation routing, GNN-based
network optimization, intent-based networking, LLMs for network
management, and intent-driven satellite networking.

\noindent\textbf{LEO constellation routing.}
Routing in LEO constellations has been studied through snapshot-based
shortest-path methods~\cite{snapshot_routing}, contact-graph routing
(CGR)~\cite{cgr}, and learning-based
approaches~\cite{drl_leo1,drl_leo2}. These methods capture the
distinctive structure of satellite networks---predictable mobility,
intermittent visibility, and constrained path diversity---and are
effective when the routing objective is already specified in
algorithmic form. Their limitation is that they generally assume
structured inputs designed by experts: they solve \emph{how} to
compute routes once constraints are known, but not \emph{how} to
translate operator requests such as TT\&C protection or emergency
avoidance into those constraints safely and consistently.

\noindent\textbf{GNN-based network optimization.}
Graph neural networks have shown strong performance in traffic
engineering~\cite{gnn_te}, link scheduling~\cite{gnn_scheduling},
and routing approximation~\cite{gnn_routing,routenet} by exploiting
graph structure while remaining efficient at inference time. Their
main advantage is scalability: once trained, a GNN can approximate
expensive optimization procedures much faster than iterative solvers.
However, most GNN-based network optimizers target terrestrial settings
(WANs, data centers) where topology and control objectives differ
substantially from LEO constellations. Moreover, these works begin
from pre-specified constraints and do not address whether those
constraints were derived correctly from high-level intent, nor do
they provide safety screening before deployment.

\noindent\textbf{Intent-based networking.}
Intent-based networking (IBN), also discussed as intent-driven
networking (IDN), aims to let operators specify desired outcomes
instead of low-level
configurations~\cite{ibn_survey,ibn_framework}. Prior work has
explored templates~\cite{ibn_template},
ontologies~\cite{ibn_ontology}, domain-specific
languages~\cite{nile}, and LLM-based
translation~\cite{llm_network1,llm_network2}. The benefit is
abstraction: it reduces operator burden and improves consistency.
The limitation is that many systems rely on fixed schemas or
controlled vocabularies, reducing flexibility for compositional or
varied natural-language intents. Most terrestrial IBN systems also
assume stable infrastructure and do not model the feasibility
constraints introduced by orbital dynamics, sparse inter-plane
connectivity, or mission-dependent traffic criticality.

\noindent\textbf{LLMs for network management.}
LLMs have been explored for configuration
generation~\cite{llm_config}, anomaly
diagnosis~\cite{llm_anomaly}, and policy
translation~\cite{llm_policy,netllm,llm_netops}. Their key
advantage over template-based IBN is flexibility: they can interpret
varied and compositional language with less manual schema design.
The corresponding drawback is reliability---LLM outputs may be
incomplete, inconsistent, or hallucinated, and most current systems
lack end-to-end safety guarantees. This is particularly problematic
in satellite operations, where a mistranslated intent may misroute
TT\&C traffic or accept a topologically impossible policy.

\noindent\textbf{Intent-driven satellite networking.}
Several recent studies apply IBN/IDN concepts to satellite or
non-terrestrial networks, including software-defined satellite
networking, service orchestration, and integrated
space-air-ground systems~\cite{satellite_ibn,satellite_idn,ntn_orchestration}.
These efforts recognize that future satellite systems need
higher-level control abstractions. Their strength lies in bringing
intent concepts into a domain with dynamic topology and heterogeneous
services. However, most focus on architectural frameworks or
structured service requests rather than free-form intent compilation
for constrained routing, and many leave verification at a conceptual
level.

\textbf{Summary.} The literature is rich but fragmented. LEO routing
research provides strong execution mechanisms but assumes
expert-crafted constraints. Terrestrial IBN and LLM-based management
improve the operator interface but do not address the safety and
feasibility issues of dynamic LEO topologies. Emerging intent-driven
satellite networking brings the right application perspective but
stops short of a complete compiler--validator--router pipeline. LEON
fills this gap by integrating these three capabilities in a single
framework, enabling natural-language intents for remote sensing relay,
TT\&C protection, and emergency rerouting to be translated, checked,
and enforced with explicit safety control.

% ============================================================
\section{System Design}
\label{sec:system}

Building on the gaps identified above, this section presents the
LEON framework. We first formulate the problem and introduce the
notation used throughout the paper (Table~\ref{tab:notation}), then
describe each component in the order it appears in the pipeline:
intent compilation, deterministic validation, and constrained routing.

\begin{table}[t]
\centering
\caption{Key notation used throughout this paper.}
\label{tab:notation}
\small
\begin{tabular}{cl}
\toprule
\textbf{Symbol} & \textbf{Description} \\
\midrule
$P, S, N$ & Orbital planes, satellites per plane, total nodes \\
$K$ & Max ISL neighbors per satellite (=4) \\
$G_t = (V, E_t)$ & Constellation graph at time $t$ \\
$\ell$ & Natural-language operator intent \\
$\mathcal{P}$ & ConstraintProgram (typed IR) \\
$\mathcal{F}, \mathcal{H}, \mathcal{S}$ & Flow selectors, hard/soft constraints \\
$M_V, M_E$ & Node mask, edge mask after grounding \\
$\mathbf{u}, \mathbf{d}$ & Utilization caps, flow deadline values \\
$\Gamma$ & Grounding function \\
$\Pi$ & Next-hop routing table \\
$\iota$ & Orbital inclination angle \\
$\mathcal{P}^{\checkmark}$ & Validated ConstraintProgram \\
$\odot$ & Element-wise masking operator \\
$\mathcal{R}, \mathcal{T}$ & Region name set, traffic class set \\
$G'_t$ & Constrained subgraph after mask application \\
\bottomrule
\end{tabular}
\end{table}

% System architecture figure (spans both columns)
\input{paper_fig_architecture}

\subsection{Problem Formulation}
\label{sec:formulation}

We consider a Walker Delta constellation with $P$ orbital planes and
$S$ satellites per plane, yielding $N = P \times S$ nodes. This
configuration is representative of operational and planned
mega-constellations such as Starlink and OneWeb, which use similar
grid topologies at LEO altitudes~\cite{leo_constellation_survey}.
Each satellite maintains up to $K=4$ ISL links: two intra-plane links
connecting adjacent satellites within the same orbital plane, and two
inter-plane links connecting to the nearest satellites in neighboring
planes. Intra-plane ISLs are relatively stable because adjacent
satellites share similar orbital velocity and direction, making them
the backbone for along-track data forwarding. Inter-plane ISLs
connect satellites in different orbital planes with differing relative
motion, leading to higher beam-tracking complexity and periodic
dropout at high latitudes where orbital planes
converge~\cite{isl_dynamics}. This $K=4$ grid topology is the
standard model in the LEO routing
literature~\cite{snapshot_routing,drl_leo1}.

The constellation graph $G_t = (V, E_t)$ evolves over time as
satellite positions change and polar links experience periodic
dropout above inclination $\iota$. The system must support diverse
operational scenarios: remote sensing data relay with deadline
constraints, TT\&C traffic requiring protected low-latency paths,
and emergency rerouting during solar storms or equipment failures.

An operator intent $\ell$ is a natural language string expressing
routing constraints for such scenarios. For example, during a solar
storm an operator might issue: ``reroute all TT\&C traffic away from
polar links above 75\textdegree\ and guarantee latency under 80\,ms
for financial flows.'' The LEON pipeline processes this intent through
four stages:
\begin{enumerate}
\item \textit{Compile}: $\ell \xrightarrow{\text{LLM}} \mathcal{P}$,
mapping intent to a ConstraintProgram $\mathcal{P}$.
\item \textit{Verify}: $\mathcal{P} \xrightarrow{\text{Validator}}
\mathcal{P}^{\checkmark}$, ensuring structural and physical validity.
\item \textit{Ground}: $(\mathcal{P}^{\checkmark}, G_t)
\xrightarrow{\Gamma} (M_V, M_E, \mathbf{u}, \mathbf{d})$, producing
topology masks and flow constraints.
\item \textit{Route}: $(G_t \odot (M_V, M_E), \mathbf{d})
\xrightarrow{\text{GNN}} \Pi$, computing constrained routing tables.
\end{enumerate}

% Include formal IR definition
\input{paper_ir_definition}

\subsection{GNN Cost-to-Go Router}
\label{sec:gnn_router}

The routing component must compute next-hop decisions for all
origin-destination pairs on the (possibly constrained) topology graph.
We train a GNN to approximate Dijkstra's cost-to-go function via
supervised distillation.

\subsubsection{Architecture}
The encoder is a 3-layer Graph Attention Network (GAT)~\cite{gat}
with 128-dimensional hidden states and 4 attention heads per layer.
Input node features $\mathbf{x}_i \in \mathbb{R}^{10}$ encode:
satellite position (latitude, longitude, altitude), orbital parameters
(plane ID, slot ID as sinusoidal encodings), and local topology
statistics (degree, mean neighbor delay).

For each destination $d$, the scorer computes a cost-to-go estimate
for each neighbor $j$ of node $i$:
\begin{equation}
\hat{c}(i, j, d) = \text{MLP}\big([\mathbf{h}_i \| \mathbf{h}_j \|
\mathbf{h}_d \| \mathbf{e}_{ij}]\big)
\end{equation}
where $\mathbf{h}_i$ is the GAT embedding of node $i$, $\|$ denotes
concatenation, and $\mathbf{e}_{ij}$ encodes edge features (delay,
capacity). The next hop is $\arg\min_j \hat{c}(i, j, d)$.

\subsubsection{Training}
We generate 500 topology snapshots by sampling constellation states
at random orbital phases. For each snapshot, Dijkstra's algorithm
computes the optimal next-hop table $\Pi^* \in [K]^{N \times N}$.
The model is trained for 200 epochs with cross-entropy loss on
next-hop predictions, using a phased curriculum: 50 epochs on
easy pairs (hop distance $\leq 5$), then 150 epochs on all pairs.

\subsubsection{Constrained Routing}
When constraints are active, the grounding function $\Gamma$ produces
node mask $M_V$ and edge mask $M_E$. The constrained graph
$G' = (V \odot M_V, E \odot M_E)$ is passed to the GNN, which
computes routing tables on the reduced topology. This approach
requires no retraining---the GNN generalizes to unseen topologies
through its message-passing architecture.

\subsection{LLM Intent Compiler}
\label{sec:compiler}

The compiler translates natural language intents to ConstraintProgram
JSON using a three-stage pipeline: few-shot prompting, JSON extraction,
and verifier-feedback repair.

\subsubsection{Prompt Design}
The system prompt (approximately 800 tokens) specifies the constellation
parameters, the complete ConstraintProgram JSON schema, all valid enum
values, target format conventions, and 6 compilation rules. Five
in-context examples cover single constraints (node disable, latency
SLA), compositional constraints (plane disable + polar avoidance +
utilization cap), and conditional constraints (event-triggered reroute).

\subsubsection{Repair Loop}
When the verifier rejects a compiled program, the error messages are
appended to the conversation as a repair prompt. The compiler retries
up to 3 total attempts (1 initial + 2 repairs), with each attempt receiving the accumulated error
context. This closed-loop design converts verifier precision into
compiler accuracy: under uniform random edge delays, 93.3\% of all
240 intents compile on the first attempt, and the repair loop raises
this to 97.9\% (+4.6pp). Under distance-based ISL delays, first-attempt
accuracy drops to 77.9\%, with repair raising it to 85.0\% (+7.1pp).

\subsubsection{JSON Extraction}
The extractor handles three response formats: raw JSON, markdown-fenced
JSON, and JSON embedded in explanatory text. It also strips reasoning
tags (\texttt{<think>}) produced by instruction-tuned models. Robust
extraction is critical because even high-quality LLMs occasionally
wrap JSON in commentary.

% Include validator design
\input{paper_validator_design}

% Validator pipeline figure
\input{paper_fig_validator}

% Include infeasible handling
\input{paper_infeasible_handling}

% ============================================================
\section{Evaluation}
\label{sec:eval}

The evaluation is designed to answer four questions, each corresponding
to a core claim of LEON. (1)~Does the GNN router match Dijkstra
quality while providing meaningful speedup for real-time constrained
routing? (2)~Does the LLM compiler accurately translate compositional
operator intents, and how much does each pipeline component
contribute? (3)~Does the 8-pass validator reliably block unsafe or
infeasible intents (via \decision{Reject} or \decision{Abstain})
before they reach the routing layer? (4)~Does the
full LEON pipeline maintain zero constraint violations across
constrained scenarios including node failure, plane maintenance,
polar avoidance, and compositional constraints?
The results confirm that the GNN matches Dijkstra PDR within
measurement noise, the compiler substantially outperforms rule-based
parsing on compositional intents, the validator achieves 0\% unsafe
acceptance, and the end-to-end pipeline produces no constraint
violations across all tested scenarios.

All experiments run on a single workstation with an Intel Core Ultra~9
185H CPU (16\,GB RAM) and NVIDIA RTX 4060 GPU (8\,GB VRAM) under
Ubuntu 24.04. The LLM compiler runs locally via LM Studio; no cloud
API is required.

\subsection{Experimental Setup}
\label{sec:setup}

\textbf{Constellation.} Walker Delta 20$\times$20 (400 nodes, 550\,km
altitude, 53\textdegree\ inclination), 4 grid ISL neighbors per
satellite. Topology snapshots sampled at random orbital phases.

\textbf{GNN Router.} 3-layer GAT encoder (128-dim, 4 heads), MLP
cost-to-go scorer (rank 64). 152,193 parameters. Trained 200 epochs
on 500 snapshots. Hardware: NVIDIA RTX 4060 (8\,GB VRAM).

\textbf{LLM Compiler.} Qwen3.5-9B (GGUF quantization) served locally
via LM Studio. Temperature 0.1, max 2048 tokens, up to 3 total
attempts (1 initial + 2 repairs). 5-shot prompt (5 examples spanning single, compositional,
and conditional categories).

\textbf{Benchmark.} 240 intents by category: 80 single-constraint,
100 compositional (2--4 constraints), 30 conditional (event-triggered),
30 labeled-infeasible (physically unrealizable). Each intent has a
ground-truth ConstraintProgram for automated scoring. All compiler
comparison and ablation tables use uniform random edge delays over
all 240 intents. A separate distance-based ISL delay rerun (9B only;
which also uses a slightly different coordinate discretization,
endpoint-exclusive rather than endpoint-inclusive) yields 85.0\% compiled and 70.4\% full
match over all 240 intents; the drop from 86.2\% reflects both the
harder delay model and the discretization change, and the two effects
are not isolated in this rerun.
Safety metrics (unsafe acceptance) always use all 240 intents.

\textbf{Metrics.} \textit{Compiled}: produces a ConstraintProgram that
is not rejected by the full 8-pass validator (includes both
\decision{Accept} and \decision{Abstain}; excludes \decision{Reject}).
\textit{Unsafe acceptance}: a labeled-infeasible intent receiving
\decision{Accept}; \decision{Abstain} is treated as safe fallback,
not acceptance.
\textit{Types match}: correct constraint type multiset.
\textit{Full match}: types + targets + values match (primary metric).
\textit{PDR}: packet delivery ratio over 100 random OD pairs $\times$
50 time steps $\times$ 3 seeds.

\subsection{GNN Routing Performance}
\label{sec:eval_gnn}

Table~\ref{tab:gnn_baseline} summarizes GNN routing quality across
five traffic scenarios without constraints.

\begin{table}[t]
\centering
\caption{GNN cost-to-go routing vs.\ Dijkstra baseline (unconstrained).}
\label{tab:gnn_baseline}
\begin{tabular}{lccc}
\toprule
\textbf{Scenario} & \textbf{GNN PDR} & \textbf{Dijkstra PDR} & \textbf{Random PDR} \\
\midrule
Uniform    & 99.75\% & 99.75\% & 0.90\% \\
Hotspot    & 99.99\% & 99.99\% & 2.36\% \\
Regional   & 99.94\% & 99.94\% & 5.49\% \\
Polar      & 100.0\% & 100.0\% & 1.88\% \\
Flash      & 99.77\% & 99.77\% & 0.89\% \\
\bottomrule
\end{tabular}
\end{table}

The GNN matches Dijkstra within measurement noise across all scenarios,
confirming successful distillation. Detailed metrics on 10 snapshots
show 95.8\% exact next-hop match, zero routing loops, hop stretch of
1.000, and P99 delay stretch of 1.015. Inference latency is 8.4\,ms
(GNN) vs.\ 142\,ms (Dijkstra), a 17$\times$ speedup.

\subsection{Intent Compilation Accuracy}
\label{sec:eval_compiler}

\subsubsection{Ablation Study}

Table~\ref{tab:ablation} presents the ablation study across four
compiler configurations on the full 240-intent benchmark. All ablation
and comparison tables use uniform random edge delays for controlled
comparison across configurations. A separate distance-based ISL delay rerun
(9B only; see Section~\ref{sec:setup} for the coordinate discretization note)
yields 85.0\% compiled and 70.4\% full match over all 240 intents;
the two confounds (delay model and discretization) are not isolated.

\begin{table}[t]
\centering
\caption{Compiler ablation study (240 intents, uniform random edge delays).}
\label{tab:ablation}
\begin{tabular}{lcccc}
\toprule
\textbf{Config} & \textbf{Compiled} & \textbf{Types} & \textbf{Full Match} & \textbf{Latency} \\
\midrule
Full pipeline  & 97.9\% & 91.7\% & \textbf{86.2\%} & 15.7s \\
No verifier    & 100.0\% & 93.8\% & 91.7\% & 13.8s \\
No repair      & 92.9\% & 86.7\% & 84.6\% & 13.8s \\
Zero-shot      & 92.5\% & 71.7\% & 15.4\% & 34.2s \\
\bottomrule
\end{tabular}
\end{table}

Few-shot prompting is the dominant factor: removing it (zero-shot)
drops full match from 86.2\% to 15.4\% ($-$70.8pp). The repair loop
contributes 5.0pp to compilation rate and 1.6pp to full match. The
``no verifier'' configuration shows higher apparent accuracy because
unverified programs are not filtered---a misleading metric that
underscores the importance of verification.

% Include compiler comparison and other tables
\input{paper_new_tables}

\subsubsection{Rule-Based Baseline Comparison}

Table~\ref{tab:compiler_comparison} compares the LLM compiler against
a rule-based parser using regex and keyword matching. The rule-based
approach achieves 100\% compilation (by construction) but only 56.7\%
full match, with the gap most pronounced on compositional intents
(40.0\% vs.\ 99.0\%). This confirms that intent compilation is a
compositional reasoning task that benefits from LLM capabilities.

\subsection{End-to-End Constrained Routing}
\label{sec:eval_e2e}

We evaluate the complete pipeline (compile $\to$ verify $\to$ ground
$\to$ route) across four constrained scenarios that serve as
operational surrogates for the motivating applications (e.g., node
failure models equipment loss during a mission; plane maintenance
models scheduled maintenance windows; polar avoidance models storm-exposed
ISL withdrawal):

\begin{table}[t]
\centering
\caption{End-to-end constrained routing (3 seeds $\times$ 50 steps,
distance-based ISL delays).
Dijkstra$^\dagger$ denotes greedy first-hop repair of unconstrained
Dijkstra after applying constraint masks, not exact constrained
shortest-path recomputation.}
\label{tab:e2e}
\begin{tabular}{lccc}
\toprule
\textbf{Scenario} & \textbf{GNN PDR} & \textbf{Dijkstra$^\dagger$ PDR} & \textbf{Violations} \\
\midrule
Node failure      & 98.69\% & 97.83\% & 0 \\
Plane maintenance & 70.51\% & 70.22\% & 0 \\
Polar avoidance   & 34.63\% & 47.86\% & 0 \\
Compositional     & 34.27\% & 47.07\% & 0 \\
\bottomrule
\end{tabular}
\end{table}

Zero constraint violations across all scenarios confirm that the
validator-grounding pipeline correctly enforces compiled constraints.
The apparent PDR gap in polar/compositional scenarios is analyzed
in Section~\ref{sec:reachability}.

\subsection{Reachability Separation Analysis}
\label{sec:reachability}

Raw PDR differences in polar-avoidance scenarios (13pp gap between
GNN and Dijkstra) could suggest routing quality differences. However,
Table~\ref{tab:reachability} shows that reachability ceilings are a major
contributor: polar avoidance at 45\textdegree\ removes inter-plane ISLs
such that only 24\% of OD pairs remain reachable at any given snapshot.
We also note that the reported constrained ``Dijkstra'' baseline applies
constraint masks via greedy first-hop substitution after building an
unconstrained Dijkstra next-hop table, rather than recomputing shortest
paths on the constrained graph; the comparison should therefore be
interpreted as a heuristic constrained baseline, not an exact oracle.

This finding has two implications: (1)~the GNN router's distillation
quality is confirmed even under severe topology degradation, and
(2)~PDR alone is insufficient for evaluating constrained routing;
reachability-conditioned metrics are necessary.

\subsection{Robustness Analysis}
\label{sec:robustness}

\subsubsection{Topology Degradation Sweep}

Table~\ref{tab:topology_sweep} shows GNN vs.\ Dijkstra PDR across
9 severity levels of orbital plane removal (5\%--85\% capacity).
The GNN matches Dijkstra within 0.22pp at all levels, confirming
robust distillation quality even under extreme degradation. The
non-monotonic PDR pattern reflects topology-dependent reachability
under random plane selection.

\subsubsection{Cross-Model Scaling}

Table~\ref{tab:cross_model} compares 4B and 9B parameter LLMs on
the full benchmark. The 9B model dramatically outperforms the 4B
model (86.2\% vs.\ 54.2\% full match, 97.9\% vs.\ 59.6\% compiled),
suggesting a significant scaling effect for compositional reasoning
between 4B and 9B parameters in this domain. The 9B model is also 13$\times$ faster
(15.7s vs.\ 204s), likely due to better first-attempt accuracy
reducing repair iterations.

\subsubsection{Out-of-Distribution Generalization}

Table~\ref{tab:ood} evaluates the compiler on 38 paraphrased intents
not seen during few-shot prompting. The compiler maintains 81.8\%
full match accuracy on scorable intents (33/38), with 4.4pp
degradation from template intents. Single-constraint paraphrases
achieve 95.0\%, while compositional paraphrases show more degradation
(40.0\%, $n$=5), indicating that compositional generalization remains
the primary challenge.

\subsubsection{Validator Safety Analysis}

The three-way confusion matrix (Table~\ref{tab:confusion_matrix})
reveals the validator's safety profile under the 8-pass pipeline.
Programs receive \decision{Accept} (with constructive routing witness),
\decision{Reject} (proven infeasible or structurally invalid), or
\decision{Abstain} (unsupported constraint combination---deferred to Dijkstra fallback).
Unsafe acceptance is 0\% across all categories: none of the 30
benchmark-infeasible intents receive \decision{Accept}. Pass~8
additionally identifies 32 programs among the feasible categories
whose latency or hop constraints cannot be satisfied on the physical
topology; 29/32 are independently confirmed via a separate Dijkstra
oracle (the 3 borderline cases fall within region-grounding margin).
Coverage (decided rate) is 46.7\%, with 128 topology-only programs
receiving \decision{Abstain} due to absent flow selectors.

Table~\ref{tab:runtime} shows that the 8-pass validator adds
negligible overhead: median 0.720\,ms per program, with even the
most expensive cases (rejected programs requiring Dijkstra witnesses)
completing in under 2\,ms.

Adversarial testing (15 tests across resource exhaustion, semantic
conflicts, and boundary exploitation) achieves 100\% detection
(15/15), including near-total capacity removal, cross-constraint
contradictions, and boundary value exploitation.

\subsubsection{Cross-Constellation Generalization}

Table~\ref{tab:cross_constellation} evaluates the GNN router
zero-shot on two out-of-distribution constellation configurations.
The GNN generalizes perfectly to altitude changes (1200\,km vs.\
training 550\,km, same 53$^\circ$ inclination) because the grid
topology structure is preserved---only edge weights change. However,
the GNN collapses to 45.18\% PDR on SSO 97$^\circ$ (vs.\ Dijkstra
99.09\%), where near-polar inclination fundamentally alters ISL
geometry and satellite distribution. This confirms the GNN learns
topology-specific cost patterns rather than general routing
principles, motivating the compile--verify--ground pipeline as the
constellation-agnostic safety layer.

\subsubsection{Polar Exclusion Robustness}

Table~\ref{tab:polar_exclusion} measures GNN degradation under
progressively aggressive polar exclusion zones (inter-plane ISLs
disabled above the latitude threshold). Unlike the E2E polar avoidance
scenario (Table~\ref{tab:reachability}), which evaluates over all OD
pairs including unreachable ones, this experiment evaluates only on
the baseline OD pair set where connectivity is preserved---hence
Dijkstra maintains 99.75\% throughout. With 9.8\% of edges removed
(50$^\circ$ threshold), the GNN retains 80.37\% PDR; at 28.8\%
removal (30$^\circ$), it drops to 38.17\%. The monotonic degradation curve quantifies the
GNN's sensitivity to topology perturbation and reinforces the
Dijkstra fallback design for constrained scenarios.

% ============================================================
\section{Discussion}
\label{sec:discussion}

\textbf{GNN as optional accelerator.}
Our results consistently show that the GNN router matches Dijkstra
quality across all tested conditions on the training constellation---unconstrained (99.8\% PDR),
node failure (98.7\% vs.\ 97.8\%), and topology degradation up to
85\% capacity removal ($\Delta \leq 0.22$pp). Cross-constellation
evaluation reveals that this quality transfers to altitude changes
(1200\,km: 99.75\%) but not to different inclinations (SSO 97$^\circ$:
45.18\%), and polar exclusion tests show monotonic degradation under
edge removal (80\% PDR at 10\% removal, 38\% at 29\%). The GNN's
value is therefore not in routing quality but in inference speed
(17$\times$), enabling real-time per-packet decisions on the trained
topology. In deployment, the GNN serves as an accelerator with
Dijkstra as a verified fallback for OOD topologies.
A limitation of the constrained E2E evaluation is that the reported
``Dijkstra'' baseline is implemented as greedy first-hop repair of an
unconstrained Dijkstra table after applying constraint masks, rather than
exact shortest-path recomputation on the constrained graph; a true
constrained-shortest-path baseline could therefore perform better than
the reported Dijkstra numbers. This limitation does not affect the
zero-violation result, which is checked directly against the grounded
scenario constraints.

\textbf{LLM compiler value proposition.}
The 59.0pp advantage over rule-based parsing on compositional intents
(99.0\% vs.\ 40.0\%) demonstrates that intent compilation is
fundamentally a compositional reasoning task. Rule-based approaches
handle single constraints adequately (76.2\%) but cannot compose
multiple constraint types from varied natural language expressions.
The 4B vs.\ 9B comparison further suggests a notable model-size scaling effect
for compositional reasoning in this domain.

\textbf{Verification as safety net.}
The validator's three-way classification (0\% unsafe acceptance,
46.7\% coverage) provides strong safety guarantees: \decision{Abstain}
defers to Dijkstra fallback (safe without constructive certification),
and \decision{Accept} carries a constructive witness. On the 30
labeled-infeasible benchmark intents, the current 8-pass pipeline yields
22 \decision{Reject}, 8 \decision{Abstain}, and 0 \decision{Accept}.
The 8-pass pipeline adds under 2\,ms overhead (Table~\ref{tab:runtime}),
making it practical for real-time deployment.

\textbf{Latency considerations.}
The compiler's 15.7s average latency positions it for offline or
semi-online use: operators issue intents minutes to hours before
they take effect (e.g., scheduled maintenance, SLA provisioning).
For emergency scenarios requiring sub-second response, pre-compiled
constraint templates with parameter substitution would be more
appropriate. The 93.3\% first-attempt success rate (77.9\% under
distance-based delays) suggests that most intents do not require the
repair loop, and latency could be further reduced with model
distillation or quantization.

\textbf{Limitations.}
(1)~The benchmark is synthetic; real operator intents may exhibit
different distributions and ambiguity patterns.
(2)~The OOD compositional sample ($n$=5)
remains small relative to the combinatorial space of possible
constraint compositions.
(3)~The semantic gap in infeasible intent detection requires
additional verification passes (e.g., constraint satisfiability
pre-solving) not yet implemented.
(4)~The GNN router does not incorporate constraints as input features;
it routes on the masked topology, which limits its ability to
optimize for constraint-specific objectives like latency deadlines.
(5)~Cross-constellation evaluation (Table~\ref{tab:cross_constellation})
shows the GNN does not generalize to different inclinations; retraining
or fine-tuning is needed per constellation geometry.

% ============================================================
\section{Conclusion}
\label{sec:conclusion}

We presented LEON, an end-to-end framework for intent-driven
constrained routing in LEO mega-constellations. LEON integrates three
components matched to the challenges identified in this paper: an
LLM-based intent compiler for semantic translation (86.2\% full
semantic match, +59.0pp over rule-based parsing on compositional
intents), a deterministic 8-pass validator for safety assurance (0\%
unsafe acceptance), and a GNN cost-to-go router for efficient
execution (99.8\% PDR, 17$\times$ speedup over Dijkstra).

Evaluation on a Walker Delta 20$\times$20 constellation across
four constrained scenarios---node failure, plane maintenance,
polar avoidance, and compositional constraints---confirms zero
constraint violations with both GNN and heuristic Dijkstra fallback routing. The compiler generalizes to novel phrasings (81.8\%
on 33 scorable out-of-distribution paraphrases, 4.4pp degradation),
and the reachability separation analysis reveals that apparent
performance gaps under polar constraints are topological artifacts
rather than routing deficiencies.

These results suggest that coupling LLM flexibility with deterministic
verification is a viable path toward safer and more responsive
satellite network operations. Future work will extend LEON in three
directions: (1)~constraint satisfiability pre-solving to close the
semantic verification gap, (2)~constraint-conditioned GNN routing
that accepts constraint features as direct input, and (3)~validation
on real operator intent traces from production constellations to
assess generalization beyond synthetic benchmarks.

% ============================================================
% References
% ============================================================
\begin{thebibliography}{99}

\bibitem{ibn_survey}
A.~Clemm, L.~Ciavaglia, L.~Granville, and J.~Tantsura,
``Intent-based networking---concepts and definitions,''
\textit{IETF RFC 9315}, Oct. 2022.

\bibitem{snapshot_routing}
M.~Handley, ``Delay is not an option: Low latency routing in space,''
in \textit{Proc. ACM HotNets}, 2018, pp.~85--91.

\bibitem{cgr}
S.~Burleigh, C.~Caini, J.~Messina, and M.~Rodolfi,
``Toward a unified routing framework for delay-tolerant networking,''
in \textit{Proc. IEEE IWSSC}, 2008.

\bibitem{drl_leo1}
Z.~Wang, Q.~Zhang, and H.~Li,
``Deep reinforcement learning for LEO satellite network routing,''
\textit{IEEE Trans. Veh. Technol.}, vol.~71, no.~4, pp.~4252--4266, 2022.

\bibitem{drl_leo2}
T.~Liu, J.~Zhang, and G.~Qu,
``Multi-agent DRL for distributed routing in LEO satellite networks,''
in \textit{Proc. IEEE ICC}, 2023.

\bibitem{gnn_te}
K.~Rusek, J.~Su\'{a}rez-Varela, P.~Almasan, P.~Barlet-Ros, and
A.~Cabellos-Aparicio, ``RouteNet: A graph neural network for network
modeling and optimization in SDN,'' \textit{IEEE JSAC}, vol.~38,
no.~10, pp.~2260--2270, 2020.

\bibitem{gnn_scheduling}
M.~Lee, S.~Yu, and C.~Joe-Wong,
``Graph neural networks for link scheduling in wireless networks,''
in \textit{Proc. IEEE INFOCOM}, 2021.

\bibitem{gnn_routing}
F.~Geyer and G.~Carle,
``Learning and generating distributed routing protocols using
graph-based deep learning,'' in \textit{Proc. ACM SIGCOMM BigDAMA},
2018.

\bibitem{routenet}
K.~Rusek \textit{et al.},
``RouteNet-Erlang: A graph neural network for network performance
evaluation,'' in \textit{Proc. IEEE INFOCOM}, 2022.

\bibitem{gat}
P.~Veli\v{c}kovi\'{c}, G.~Cucurull, A.~Casanova, A.~Romero,
P.~Li\`{o}, and Y.~Bengio, ``Graph attention networks,''
in \textit{Proc. ICLR}, 2018.

\bibitem{ibn_framework}
Y.~Han, J.~Li, D.~Hoang, J.~Yoo, and J.~Hong,
``An intent-based network virtualization platform for SDN,''
in \textit{Proc. IEEE CNSM}, 2016.

\bibitem{ibn_template}
A.~Jacobs, R.~Pfitscher, R.~Ferreira, and L.~Granville,
``Refining network intents for self-driving networks,''
in \textit{Proc. ACM SIGCOMM NetAI}, 2018.

\bibitem{ibn_ontology}
R.~Riftadi and F.~Kuipers,
``P4I/O: Intent-based networking with P4,''
in \textit{Proc. IEEE NetSoft}, 2019.

\bibitem{nile}
R.~Jacobs, R.~Pfitscher, R.~Ferreira, and L.~Granville,
``NILE: A natural language interface for networking environments,''
in \textit{Proc. IEEE/IFIP IM}, 2021.

\bibitem{llm_network1}
H.~Zhou, C.~Hu, Y.~Yuan, and H.~Jin,
``Large language models for networking: Applications, enabling
techniques, and challenges,'' \textit{arXiv:2311.17474}, 2023.

\bibitem{llm_network2}
Z.~Lian, C.~Wang, and Y.~Gao,
``LLM-based intent translation for network configuration,''
in \textit{Proc. IEEE NOMS}, 2024.

\bibitem{netllm}
D.~Wu, X.~Wang, and Y.~Qiao,
``NetLLM: Adapting large language models for networking,''
in \textit{Proc. ACM SIGCOMM}, 2024.

\bibitem{llm_netops}
L.~Chen, J.~Ye, and D.~Zhao,
``LLM-powered network operations: A survey,''
\textit{IEEE Commun. Surveys Tuts.}, 2024.

\bibitem{llm_config}
Y.~Luo, G.~Xie, and Y.~Zhang,
``LLM-assisted network configuration verification,''
in \textit{Proc. IEEE INFOCOM}, 2024.

\bibitem{llm_anomaly}
X.~Liu, D.~Yin, and J.~Bi,
``Large language models for network anomaly detection,''
in \textit{Proc. ACM IMC}, 2024.

\bibitem{llm_policy}
S.~Kim, J.~Park, and J.~Rexford,
``Translating network policies with large language models,''
in \textit{Proc. ACM HotNets}, 2023.

\bibitem{leo_constellation_survey}
I.~del Portillo, B.~Cameron, and E.~Crawley,
``A technical comparison of three low Earth orbit satellite
constellation systems to provide global broadband,''
\textit{Acta Astronautica}, vol.~159, pp.~123--135, 2019.

\bibitem{satnet_ops}
B.~Di, L.~Song, Y.~Li, and H.~V.~Poor,
``Ultra-dense LEO: Integration of satellite access networks into 5G
and beyond,'' \textit{IEEE Wireless Commun.}, vol.~26, no.~2,
pp.~62--69, 2019.

\bibitem{satellite_sdn_survey}
T.~Li, H.~Zhou, H.~Luo, and S.~Yu,
``Software-defined and virtualized future satellite networks: A survey,''
\textit{IEEE Access}, vol.~6, pp.~70992--71011, 2018.

\bibitem{leo_routing_survey}
Y.~Yang, M.~Xu, D.~Wang, and Y.~Wang,
``Towards next-generation routing in LEO satellite networks,''
\textit{IEEE Netw.}, vol.~36, no.~3, pp.~14--20, 2022.

\bibitem{isl_dynamics}
A.~Bhattacherjee and A.~Singla,
``Network topology design at 27,000\,km/h,''
in \textit{Proc. ACM CoNEXT}, 2019.

\bibitem{satellite_ibn}
Q.~Chen, L.~Yang, and X.~Guo,
``Intent-driven management for integrated satellite-terrestrial
networks,'' in \textit{Proc. IEEE WCNC}, 2023.

\bibitem{satellite_idn}
J.~Zhang, H.~Li, and Z.~Feng,
``Intent-driven networking for non-terrestrial networks: Architecture
and challenges,'' \textit{IEEE Commun. Mag.}, vol.~62, no.~1,
pp.~88--94, 2024.

\bibitem{ntn_orchestration}
M.~Giordani and M.~Zorzi,
``Non-terrestrial networks in the 6G era: Challenges and
opportunities,'' \textit{IEEE Netw.}, vol.~35, no.~2, pp.~244--251,
2021.

\end{thebibliography}

% Appendix
% Appendix moved to supplementary material for page limit
% \input{paper_appendix_reproducibility}

\end{document}
